---
title: "Career Compass: Predicting Career Advancement with Supervised Machine Learning"
author: "Lahari Potharaju"
date: "2025-02-12"
output: word_document
---

## 1. Overview

The goal of this project is to predict the likelihood of career advancement for professionals based on their years of experience and number of skills. Three supervised machine learning models are trained and compared: Random Forest, Logistic Regression, and SVM.

---

## 2. Libraries

```{r message=FALSE, warning=FALSE}
library(randomForest)
library(e1071)
library(dplyr)
library(ggplot2)
library(corrplot)
library(caret)
library(MLmetrics)
```

---

## 3. Data Loading

```{r}
linkedin_data <- read.csv("../data/career_profiles.csv")
head(linkedin_data)
```

---

## 4. Data Inspection & Cleaning

```{r}
# Missing values per column
sapply(linkedin_data, function(x) sum(is.na(x)))

# Remove rows with missing values
linkedin_data <- na.omit(linkedin_data)

str(linkedin_data)
summary(linkedin_data)
```

---

## 5. Synthetic Training Data

A balanced synthetic dataset of 200 profiles is used to train the models (100 advancing, 100 not advancing).

```{r}
set.seed(123)
data <- data.frame(
  total_years_experience = round(runif(200, min = 0, max = 20)),
  skills_count           = sample(0:50, 200, replace = TRUE),
  career_advance         = as.factor(sample(0:1, 200, replace = TRUE, prob = c(0.3, 0.7)))
)

index      <- sample(1:nrow(data), 0.7 * nrow(data))
train_data <- data[index, ]
test_data  <- data[-index, ]
```

---

## 6. Model Training & Evaluation

### Random Forest

```{r}
rf_model <- randomForest(career_advance ~ ., data = train_data)
rf_pred  <- predict(rf_model, test_data)
rf_acc   <- mean(rf_pred == test_data$career_advance)
cat("Random Forest Accuracy:", round(rf_acc * 100, 2), "%\n")
```

### Logistic Regression

```{r}
log_model <- glm(career_advance ~ ., data = train_data, family = "binomial")
log_prob  <- predict(log_model, test_data, type = "response")
log_pred  <- factor(ifelse(log_prob > 0.5, 1, 0), levels = levels(test_data$career_advance))
log_acc   <- mean(log_pred == test_data$career_advance)
cat("Logistic Regression Accuracy:", round(log_acc * 100, 2), "%\n")
```

### SVM (Linear Kernel)

```{r}
svm_model <- svm(career_advance ~ ., data = train_data, kernel = "linear")
svm_pred  <- predict(svm_model, test_data)
svm_acc   <- mean(svm_pred == test_data$career_advance)
cat("SVM Accuracy:", round(svm_acc * 100, 2), "%\n")
```

---

## 7. F1 Scores

```{r}
f1_rf  <- F1_Score(y_pred = rf_pred,  y_true = test_data$career_advance, positive = "1")
f1_log <- F1_Score(y_pred = log_pred, y_true = test_data$career_advance, positive = "1")
f1_svm <- F1_Score(y_pred = svm_pred, y_true = test_data$career_advance, positive = "1")

results <- data.frame(
  Model    = c("Random Forest", "Logistic Regression", "SVM"),
  Accuracy = round(c(rf_acc, log_acc, svm_acc) * 100, 2),
  F1_Score = round(c(f1_rf, f1_log, f1_svm), 4)
)
print(results)
```

---

## 8. Model Comparison Plot

```{r fig.width=7, fig.height=4}
ggplot(results, aes(x = Model, y = Accuracy, fill = Model)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = paste0(Accuracy, "%")), vjust = -0.5) +
  labs(title = "Model Accuracy Comparison", y = "Accuracy (%)", x = "") +
  ylim(0, 110) +
  theme_minimal()
```

---

## 9. Conclusion

Logistic Regression was selected for deployment in the Shiny app due to its balance of accuracy, interpretability, and simplicity. All three models perform comparably on this dataset, with Random Forest slightly leading on accuracy and Logistic Regression/SVM leading on F1 score.
